---
title: "Covid91 vaccine study Final2022"
author: "Iven Santos"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final_Data.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
Males = filter(FinalData, sex == "M")
Females = filter(FinalData, sex == "F")
LGBTQ = filter(FinalData, LGBTQ == "gay")
DrugUser = filter(FinalData, DrugUser == "yes")
```
# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of sub-populations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Null Hypothesis

$H_0$: There's no shown difference between the vaccine and the group it gets used on. Both values are independent.

$H_A$: There is a shown difference between the vaccine and the group it gets used on. Both values are dependent.

## Males

### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment, data = Males)
barchartGC(~infected + treatment, data = Males, type = "percent")
```

This chart shows that the males who were given the vaccine are less likely to get covid19 compared to those who didn't get it.

### Numerical Descriptive Results
```{r}
table1 <- xtabs(~infected + treatment, data = Males)
rowPerc(table1)
colPerc(table1)
```

Those who are vaccinated are least likely to get sick. As the majority of those who are infected are the ones who were given the placebo. Proving that the vaccine is effective at keeping males from getting infected.

### Inferential Descriptive Results 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
This table continues to confirms that the vaccine is good at keeping people healthy as there is a 128 person difference in the infections.


### Male Efficacy

```{r}
efficMales = 2.09/(2.09 + 1)*100
efficMales = round(efficMales)
confIntLower = 1.67/(1.67+1) * 100 
confIntLower = round(confIntLower)
confIntUpper = 2.64/(2.64 + 1 ) * 100 
confIntUpper = round(confIntUpper)
```

The vaccine is shown multiple times to be effective on males. 

## Females

### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment, data = Females)
barchartGC(~infected + treatment, data = Females, type = "percent")
```

As with the chart for males, this also appears to show that women who have the covid19 vaccine are also less susceptible to contracting it.

### Numerical Descriptive Results
```{r}
table2 <- xtabs(~infected + treatment, data = Females)
rowPerc(table2)
colPerc(table2)
```
The table further shows that giving women the vaccine makes them less likely to get sick from covid19 and are more vulnerable when taking the placebo.

### Inferential Descriptive Results 
```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

We reject the null hypothesis on all three tables as they show that there is benefit for women to have the vaccine.

### Female  Efficacy 

```{r}
efficFemales = 1.43/(1.43 + 1)*100
efficFemales = round(efficFemales)
confIntLower = 1.15/(1.15 + 1) * 100 
confIntLower = round(confIntLower)
confIntUpper = 1.78/(1.78 + 1 ) * 100 
confIntUpper = round(confIntUpper)
```

The vaccine is shown three times here that it is effective.

## LGBTQ

### Graphical Descriptive Results 

```{r}
barchartGC(~infected + treatment, data = LGBTQ)
barchartGC(~infected + treatment, data = LGBTQ, type = "percent")
```

The chart here shows that the vaccine actually makes the LGBTQ more likely to catch covid19 than if they were with out it.

### Numerical Descriptive Results

```{r}
table3 <- xtabs(~infected + treatment, data = LGBTQ)
rowPerc(table3)
colPerc(table3)
```

This table further proves the first chart as about 86.% of people sick are ones who have taken the covid19 vaccine.

### Inferential Descriptive Results
```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```
All three of these tables show that getting the vaccine as a person of LGTBQ has more risks than it does benefits. Which in turn disproves the null hypothesis.

### LGBTQ Efficacy 

```{r}
efficLGBTQ = .15/(.15 + 1)*100
efficLGBTQ = round(efficLGBTQ)
confIntLower = .06/(.06 + 1) * 100 
confIntLower = round(confIntLower)
confIntUpper = .31/(.31 + 1 ) * 100 
confIntUpper = round(confIntUpper)
```

Having the vaccine has negative effects for those of LGBTQ

## Drug Users

### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment, data = DrugUser)
barchartGC(~infected + treatment, data = DrugUser, type= "percent")
```

Drug users who take the vaccine become noticeably more vulnerable to getting infected when they have the vaccine.

### Numerical Descriptive Results 
```{r}
table4 <- xtabs(~infected + treatment, data = DrugUser)
rowPerc(table4)
colPerc(table4)
```
85% of the drug users who are infected are the ones who are vaccinated. Showing even further that the vaccine is doing them more harm.

### Inferential Descriptive Results 
```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```

The tables shows that by having the vaccine that the drug users put themselves more at risk for getting sick. So the null hypothesis has to be rejected.

### Drug User Efficacy 

```{r}
efficDrugUsers = .18/(.18 + 1)*100
efficDrugUsers = round(efficDrugUsers)
confIntLower = .08/(.08 + 1) * 100 
confIntLower = round(confIntLower)
confIntUpper = .37/(.37 + 1 ) * 100 
confIntUpper = round(confIntUpper)
```

Taking the vaccine makes drug users liklier to get infected with covid19.

## Conclusion

The original question was if the vaccine has the same effect on every user or if there is a difference. And the answer is, there's a difference. While for Males and Females it proved that taking the vaccine was beneficial for them in protecting against covid19. Other charts showed the exact opposite for LGBTQ members and Drug Users, who were shown multiple times to actually have negative results when taking the vaccine, that they were more likely to get sick from it.